---
output:
  html_document:
    df_print: paged
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.width=10, fig.height=6, fig.align = "left") 
```



```{r }

library(tidyverse)
library(devtools)
library(ggmap)
library(RColorBrewer)
library(colorspace)
library(here)
library(sp)
library(plotly)
library(leaflet)
library(birdnames)

options(scipen = 999)
source("C:/Users/scott.jennings/Documents/Projects/my_R_general/utility_functions/numbers2words.R")
source("C:/Users/scott.jennings/Documents/Projects/core_monitoring_research/HEP/HEP_data_work/HEP_code/HEP_utility_functions.R")
# source_url("https://raw.githubusercontent.com/scottfjennings/HEP_data_work/master/HEP_code/HEP_utility_functions.R")


zyear = 2021
```



```{r }
#  data ----
hep_sites <- readRDS(here("HEP_data/HEP_site_names_nums_utm")) %>% 
  filter(!is.na(utmeast))

hep_sites_utm <- SpatialPointsDataFrame(coords = cbind(hep_sites$utmeast, hep_sites$utmnorth), data = hep_sites, proj4string = CRS("+proj=utm +zone=10 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")) %>% 
  spTransform(CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))

# total nests counted across the study area

hep_abund <- readRDS(here("HEP_data/hep_annual_nest_abundance")) %>% 
  mutate(spp.name = translate_bird_names(species, "alpha.code", "common.name"))

hep_abund_study_area <- hep_abund %>% 
  cut_never_nested() %>% 
  filter(peakactvnsts > 0, year >= 1990) %>% 
  group_by(year, species, spp.name) %>% 
  summarise(tot.abund = sum(peakactvnsts),
            num.colonies = n())



hep_changes <- readRDS(here("HEP_data/colony_changes_bycode")) %>% 
  mutate(spp.name = translate_bird_names(species, "alpha.code", "common.name"))


# annual mean percent change in colony size
ann_change_study_area <- hep_changes %>% 
  filter(!is.na(per.change), !colony.status %in% c("stay.abandoned", "colonized"), year >= 1990) %>% 
  group_by(year, species, spp.name) %>% 
  summarise(num.colonies = n(),
            mean.per.change = mean(per.change)) %>% 
  ungroup()
         

# 

# probably not using weigted percent change, keeping this code here for now  
#  mutate(tot.nests = sum(peakactvnsts),       
#         colony.weight = peakactvnsts/tot.nests,
#         weighted.abs.change = abs.change.1year * colony.weight,
#         weighted.per.change = per.change.1year * colony.weight) 


#ann_change_study_area <- ann_change_study_area1 %>% 
#  filter(zero2some == FALSE) %>% 
#  summarise(year.weighted.abs.change = sum(weighted.abs.change*is.finite(weighted.abs.change),na.rm=TRUE),
#         per.change.1year = sum(weighted.per.change*is.finite(weighted.per.change),na.rm=TRUE),
#         num.colonies = n(),
#         num.nests = sum(peakactvnsts),
#         mean.per.change = mean(per.change.1year),
#         absolute.change = sum(abs.change.1year)) %>% 
#  ungroup()  
#  pivot_longer(cols = contains("change"), names_to = "type.change", values_to = "amount.change") %>% 
#  mutate(change.group = case_when(type.change == "absolute.change" ~ "Absolute change in raw\nnumber of nests",
#                                  type.change == "mean.per.change" ~ "Mean percent change\nin number of nests",
#                                  type.change == "year.weighted.abs.change" ~ "Absolute change in raw\nnumber of nests\n(weighted by colony size)",
#                                  type.change == "per.change.1year" ~ "Mean percent change\nin number of nests\n(weighted by colony size)"))
```


```{r}
# function to make the annual change figure
annual_abund_plotter <- function(zspecies) {
  
z_name_color = filter(spp_color_name, species == zspecies)

ann_abund <- hep_abund_study_area  %>% 
  filter(species == zspecies)

y2020 <- filter(ann_abund, year == 2020)$tot.abund * 0.6
  
ztitle <- paste("Average % change in", z_name_color$spp.name, "colony size.")
max_plot_year <- ceiling(max(ann_abund$year)/5) * 5
  
ann_abund_plot <- ann_abund %>%  
ggplot() +
  geom_point(aes(year, tot.abund,
                 text = paste(year,  "\n",
                              "# monitored nests: ", tot.abund, "\n",
                              "# monitored colonies: ", num.colonies, "\n")), color = z_name_color$spp.color) +
  stat_smooth(aes(year, tot.abund), color = z_name_color$spp.color, se = FALSE) +
  annotate("text", x = 2020, y = y2020, label = "Reduced\neffort\nin 2020") +
  theme_bw() +
  theme(legend.position = "none") +
  ylab("") +
  #ggtitle(ztitle) +
  scale_x_continuous(breaks = seq(1990, max_plot_year, by = 5))

ggplotly(ann_abund_plot, tooltip = "text")
}

```


```{r}
# function to make the annual change figure
annual_change_plotter <- function(zspecies) {
  
z_name_color = filter(spp_color_name, species == zspecies)

ann_change <- ann_change_study_area  %>% 
  filter(species == zspecies)

y2020 <- min(ann_change$mean.per.change) * 1.1
  
ztitle <- paste("Average % change in", z_name_color$spp.name, "colony size.")
max_plot_year <- ceiling(max(ann_change$year)/5) * 5
  
ann_change_plot <- ann_change %>%  
ggplot() +
  #geom_rect(aes(xmin = min(ann_change$year), xmax = max(ann_change$year), ymin = -100, ymax = 0), fill = brewer.pal(9, "Reds")[7], alpha = 0.2) +
  #geom_rect(aes(xmin = min(ann_change$year), xmax = max(ann_change$year), ymin = 0, ymax = max(ann_change$mean.per.change)*1.2), fill = brewer.pal(9, "Blues")[8], alpha = 0.2) +
  #annotate("text", x = max(ann_change$year)-4, y = max(ann_change$mean.per.change), label = "Population increasing") +
  #annotate("text", x = max(ann_change$year)-4, y = -90, label = "Population decreasing") +
  geom_point(aes(year, mean.per.change,
                 text = paste(year,  "\n",
                              #"# monitored nests: ", tot.nests, "\n",
                              #"# monitored colonies: ", num.colonies, "\n",
                              "Average % change from", year - 1, ":", round(mean.per.change, 1))), color = z_name_color$spp.color) +
  stat_smooth(aes(year, mean.per.change), color = z_name_color$spp.color, se = FALSE) +
  geom_hline(yintercept = 0) +
  theme_bw() +
  theme(legend.position = "none") +
  ylab("") +
  annotate("text", x = 2020, y = y2020, label = "Reduced\neffort\nin 2020") +
  ylim(min(ann_change$mean.per.change) * 1.3, max(ann_change$mean.per.change) * 1.1) +
  #ggtitle(ztitle) +
  scale_x_continuous(breaks = seq(1990, max_plot_year, by = 5))

ggplotly(ann_change_plot, tooltip = "text")
}

```



```{r} 
# function to make the map

hep_report_year_spp_mapper <- function(zspecies, zyear) {



spp_report_year <- hep_changes  %>% 
  filter(year == zyear, species == zspecies, !is.na(per.change)) %>% 
  left_join(., data.frame(hep_sites_utm))

spp_report_year_utm <- SpatialPointsDataFrame(coords = cbind(spp_report_year$utmeast, spp_report_year$utmnorth), data = spp_report_year, proj4string = CRS("+proj=utm +zone=10 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

spp_report_year_dd <- spTransform(spp_report_year_utm, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))
spp_report_year_dd_df <- data.frame(spp_report_year_dd) %>% 
  rename(lon = coords.x1, lat = coords.x2) %>% 
  mutate(hollow.fill = ifelse(colony.status == "abandoned", TRUE, FALSE)) %>% 
  filter(colony.status != "colonized")


# for centering the map:
study_center_lat <- max(spp_report_year_dd_df$lat) - ((max(spp_report_year_dd_df$lat) - min(spp_report_year_dd_df$lat))/2)
study_center_lng <- max(spp_report_year_dd_df$lon) - ((max(spp_report_year_dd_df$lon) - min(spp_report_year_dd_df$lon))/2)

# labels for popup bubbles
labs <- lapply(seq(nrow(spp_report_year_dd_df)), function(i) {
  paste0( '<p>', spp_report_year_dd_df[i, "site.name"], '<p></p>', 
          #"% change from last year: ", spp_report_year_dd_df[i, "per.change.1year"],  '<p></p>',
          "# nests in ", zyear, ": ", spp_report_year_dd_df[i, "peakactvnsts"],'</p><p>', 
          "# nests in ", zyear-1, ": ", spp_report_year_dd_df[i, "prev.yr.nsts"]) 
})


# make color scales to indicate % change
# greatest decrease
abs_low_change = ceiling(abs(min(spp_report_year_dd_df$per.change, na.rm = TRUE)))
# greatest increase
abs_high_change = ceiling(max(spp_report_year_dd_df$per.change, na.rm = TRUE))

# if there are only positive changes, only use blues
if(min(spp_report_year_dd_df$per.change, na.rm = TRUE) >= 0) {
  rampcols <- colorRampPalette(colors = c(brewer.pal(9, "Blues")[3], brewer.pal(9, "Blues")[8]), space = "Lab")(abs_high_change)
}

# if there are only negative changes, only use reds
if(max(spp_report_year_dd_df$per.change, na.rm = TRUE) <= 0) {
  rampcols <- colorRampPalette(colors = c(brewer.pal(9, "Reds")[7], brewer.pal(9, "Reds")[3]), space = "Lab")(abs_low_change)
}


# if there are positive and negative changes, use blues and reds
if(max(spp_report_year_dd_df$per.change, na.rm = TRUE) >= 0 & min(spp_report_year_dd_df$per.change, na.rm = TRUE) <= 0) {
  rampcols <- c(colorRampPalette(colors = c(brewer.pal(9, "Reds")[7], brewer.pal(9, "Reds")[3]), space = "Lab")(abs_low_change),
                colorRampPalette(colors = c(brewer.pal(9, "Blues")[3], brewer.pal(9, "Blues")[8]), space = "Lab")(abs_high_change))
}


# vector of colors for negative changes
#rc1 <- colorRampPalette(colors = c(brewer.pal(9, "Reds")[7], brewer.pal(9, "Reds")[3]), space = "Lab")(abs_low_change)

# vector of colors for positive changes
#rc2 <- colorRampPalette(colors = c(brewer.pal(9, "Blues")[3], brewer.pal(9, "Blues")[8]), space = "Lab")(abs_high_change)

## Combine the two color palettes
# rampcols <- c(rc1, rc2)

mypal <- colorNumeric(palette = rampcols, domain = spp_report_year_dd_df$per.change)


fix_radius <- function(x) {
  zradius = (log10(x + 2) + 0.1) * 10
}


spp_report_year_dd_df %>%
leaflet() %>%
addProviderTiles(providers$Stamen.TonerLite, group = "Toner Lite") %>%
addProviderTiles(providers$Esri.WorldImagery, group = "World Imagery") %>%
addLayersControl(baseGroups = c("World Imagery", "Toner Lite")) %>%
  addCircleMarkers(radius = ~fix_radius(peakactvnsts),
                   color = ~mypal(per.change),
                   opacity = 0.8,
                   fillOpacity = ~ifelse(hollow.fill == TRUE, 0, 0.8),
                   label=lapply(labs, htmltools::HTML),
                   labelOptions = labelOptions(textsize = "15px")) %>%
addMiniMap(
    toggleDisplay = TRUE,
    tiles = providers$Stamen.TonerLite
    ) %>% 
  setView(lat = study_center_lat * 1.001, # nudge up from center
          lng = study_center_lng * 0.999, # nudge left from center
          zoom = 9)%>%
  addLegend("bottomleft", pal = mypal, values = ~per.change,   opacity = 1, title = paste(zyear-1, "-", zyear, "</br>% change", sep = ""))


  

}



```


```{r}

# create reactive text blocks to be inserted into the document


prev.year = zyear-1

total_colonies_monitored <- hep_abund %>% 
  filter(year == zyear, species %in% hep_species_order[1:4,]$species) %>% 
  distinct(code) %>% 
  nrow()

intro_numbers <- hep_changes %>% 
  filter(year == zyear, peakactvnsts > 0, species %in% hep_species_order[1:4,]$species) %>% 
  group_by(species, spp.name) %>%
  summarise(num.colonies = n()) %>% 
  ungroup() %>% 
  mutate(out.num.colonies = ifelse(num.colonies < 10, text_number(num.colonies), num.colonies),
         colony.text = ifelse(num.colonies == 1, "colony", "colonies"),
         out.text = paste(spp.name, "s nested in ", out.num.colonies, sep = ""),
         spp.name = factor(spp.name, levels = levels(core4spp))) 
  
intro_numbers_text <- paste("In ", zyear, ", we monitored a total of ", total_colonies_monitored, " colonies. Of these colonies, ", 
                            filter(intro_numbers, species == "GREG")$out.text, ", ",
                            filter(intro_numbers, species == "GBHE")$out.text, ", ", 
                            filter(intro_numbers, species == "SNEG")$out.text, ", and ",
                            filter(intro_numbers, species == "BCNH")$out.text, ".", sep = "")


report_year_annual_change <- ann_change_study_area %>% 
  filter(year == zyear) %>% 
  mutate(annual.change.text = case_when(mean.per.change < 0 ~ paste("On average, the number of nests in ", spp.name, " colonies decreased from ", prev.year, " by ", round(abs(mean.per.change), 1), "%.", sep = ""),
                                  mean.per.change > 0 ~ paste("On average, the number of nests in ", spp.name, " colonies increased by ", round(abs(mean.per.change), 1), "% from ", prev.year, ".", sep = "")))

report_year_num_nests <- hep_changes %>% 
  filter(year == zyear, peakactvnsts > 0) %>% 
  group_by(species, spp.name) %>% 
  summarise(total.nests = sum(peakactvnsts),
            total.colonies = n()) %>% 
  mutate(nest.text = ifelse(total.nests == 1, "nest", "nests"),
         colony.text = ifelse(total.colonies == 1, "colony", "colonies"),
         summary.text = paste("In ", zyear, ", we counted ", text_number(total.nests), " ", spp.name, " ", nest.text, " in ", text_number(total.colonies), " active ", colony.text, ".", sep = ""))

report_year_up_down <- hep_changes %>% 
  filter(year == zyear, !is.na(change.dir)) %>% 
  group_by(species, spp.name, change.dir)  %>% 
  summarise(num.change.dir = n()) %>% 
  ungroup() %>% 
  pivot_wider(id_cols = c(spp.name, species), names_from = change.dir, values_from = num.change.dir) %>% 
  mutate(across(c("down", "stable", "up"), ~replace_na(.,0))) %>% 
  data.frame() %>% 
  mutate(up.text = ifelse(up == 1, "colony", "colonies"),
         down.text = ifelse(down == 1, "colony", "colonies"),
         stable.text = ifelse(stable == 1, "colony", "colonies"),
         change.dir.text = paste("Taking a closer look at each individual colony, in ", zyear, ", ", text_number(up), " ", spp.name, " ", up.text, " increased in size, ",
                              text_number(down), " ", down.text, " decreased, and ", 
                              text_number(stable), " ", stable.text, " remained the same size.", sep = ""))


# ---
report_year_zero2some <- hep_changes %>% 
  filter(year == zyear, !is.na(peakactvnsts), colony.status == "colonized") %>% 
  left_join(hep_sites) %>% 
  group_by(species, spp.name) %>% 
  summarise(col.names = paste(site.name, collapse="; "),
            num.cols = n()) %>% 
  mutate(col.text = ifelse(num.cols == 1, "colony", "colonies"),
         zero2some.text = paste(spp.name,  "s nested in ", text_number(num.cols), " ", col.text, " where they did not nest in ", prev.year, " (hollow blue circles).", sep = ""))
  

report_year_some2zero <- hep_changes %>% 
  filter(year == zyear, colony.status == "abandoned") %>% 
  left_join(hep_sites) %>% 
  group_by(species, spp.name) %>% 
  summarise(col.names = paste(site.name, collapse="; "),
            num.cols = n()) %>% 
  mutate(col.text = ifelse(num.cols == 1, "colony", "colonies"),
         some2zero.text = paste("There were ", text_number(num.cols), " active ", spp.name, " ", col.text, " in ", prev.year, " that had zero nests in ", zyear, " (hollow red circles).", sep = ""))


# ---
annual_change_texter <- function(zspecies) {
  paste(filter(report_year_num_nests, species == zspecies)$summary.text, filter(report_year_annual_change, species == zspecies)$annual.change.text)
}




new_colony_texter <- function(zspecies) {
  if(nrow(filter(report_year_zero2some, species == zspecies)) == 0) {
    ""
  } else {
  filter(report_year_zero2some, species == zspecies)$zero2some.text
  }
}

abandoned_colony_texter <-function(zspecies) {
  if(nrow(filter(report_year_some2zero, species == zspecies)) == 0) {
    ""
  } else {
  filter(report_year_some2zero, species == zspecies)$some2zero.text
  }
}


map_texter <- function(zspecies) {
  paste(filter(report_year_up_down, species == zspecies)$change.dir.text,
        new_colony_texter(zspecies),
        abandoned_colony_texter(zspecies),
        sep = "")
}

general_map_text <- "On this map, the size of the circle represents colony size (number of nests). You can zoom in and pan, and hover over each colony to get more information."

```

## How are the egrets doing? {.tabset .tabset-fade}
<font size="4"> 
Herons and Egrets are charismatic symbols of wetlands, and they play important roles in how wetlands and other ecosystems function. Tracking the number of herons and egrets in an area can give us a window into the health of the local environment.
</font>

<br>

<font size="4">
Every year since 1989, a dedicated crew of volunteers has fanned out over the northern San Francisco Bay Area to monitor nesting herons and egrets. They count how many nests are in each colony, and estimate the number of young produced in each nest. Long term monitoring projects like this help use understand how ecosystems change over time as humans continue to alter the landscape and climate continues to warm.
</font>

<br>

<font size="4">
`r intro_numbers_text`
</font>

<br>

<font size="4">
We can estimate year-to-year changes in the number of birds breeding across the entire region by first calculating the percent change in size of each colony, then averaging those values across all the colonies monitored. These data show that in recent years the local population of Great Egrets is increasing, the local Great Blue Heron population is generally stable, and populations of Snowy Egrets and Black-crowned Night-Herons are generally stable but somewhat more variable year-to year than the other species. You can explore the figures and maps below for more information on each species.

<br>

</font>

### Great Egret

<br>
<font size="4">
`r annual_change_texter("GREG")` Overall, Great Egret colonies have increased in size during most of this monitoring project. This suggests that the local population of Great Egrets is increasing.
</font>
 

<br>
<font size="5">
Total Great Egret nesting abundance, 1990 to `r zyear`.
</font>
```{r}

annual_abund_plotter("GREG")
```


<br>
<font size="5">
Average annual % change in Great Egret colony size, 1990 to `r zyear`.
</font>
```{r}

annual_change_plotter("GREG")
```

<br>
<font size="4">
`r map_texter("GREG")`
</font>

<br>
<font size="5">
Great Egret colonies in `r zyear`.
</font>

```{r}
hep_report_year_spp_mapper("GREG", zyear)
```




### Great Blue Heron


<br>
<font size="4">
`r annual_change_texter("GBHE")` Overall, the number of nests in Great Blue Heron colonies has increased on average over the last several years, although often not by very much. This suggests that the local Great Blue Heron population is generally stable.
</font>

 

<br>
<font size="5">
Total Great Blue Heron nesting abundance, 1990 to `r zyear`.
</font>
```{r}

annual_abund_plotter("GBHE")
```


<br>
<font size="5">
Average annual % change in Great Blue Heron colony size, 1990 to `r zyear`.
</font>
```{r}

annual_change_plotter("GBHE")
```

<br>
<font size="4">
`r map_texter("GBHE")`
</font>

<br>
<font size="5">
Great Blue Heron colonies in `r zyear`.
</font>

```{r}
hep_report_year_spp_mapper("GBHE", zyear)
```

### Snowy Egret


<br>
<font size="4">
`r annual_change_texter("SNEG")`The number of nests in Snowy Egret colonies has generally been stable during our monitoring project, suggesting an overall stable population. However, there have been occasional single year spikes in colony growth, which we don't yet fully understand.
</font>
 

<br>
<font size="5">
Total Snowy Egret nesting abundance, 1990 to `r zyear`.
</font>
```{r}

annual_abund_plotter("SNEG")
```


<br>
<font size="5">
Average annual % change in Snowy Egret colony size, 1990 to `r zyear`.
</font>
```{r}

annual_change_plotter("SNEG")
```

<br>
<font size="4">
`r map_texter("SNEG")`
</font>

<br>
<font size="5">
Snowy Egret colonies in `r zyear`.
</font>

```{r}
hep_report_year_spp_mapper("SNEG", zyear)
```

### Black-crowned Night-Heron
 



<br>
<font size="4">
`r annual_change_texter("BCNH")` In recent years, the number of nests in Black-crowned Night-Heron colonies has tended to increase and decrease more than the other species we monitor. This means that although the population appears generally stable over the long term, there can be dramatic changes in the number of nesting birds from one year to the next.
</font>

<br>
<font size="5">
Total Black-crowned Night-Heron nesting abundance, 1990 to `r zyear`.
</font>
```{r}

annual_abund_plotter("BCNH")
```


<br>
<font size="5">
Average annual % change in Black-crowned Night-Heron colony size, 1990 to `r zyear`.
</font>
```{r}

annual_change_plotter("BCNH")
```

<br>
<font size="4">
`r map_texter("BCNH")`
</font>

<br>
<font size="5">
Black-crowned Night-Heron colonies in `r zyear`.
</font>

```{r}
hep_report_year_spp_mapper("BCNH", zyear)
```

<br>

##
<font size="4">
`r general_map_text`
</font>
