---
title: ''
output:
  word_document: default
  html_document:
    df_print: paged
date: "`r Sys.Date()`"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      fig.width = 7,
                      fig.height = 5)
```



```{r}
library(tidyverse)
library(cowplot)
library(gtable)
library(grid)
library(MASS)
library(birdnames)
library(here)
library(flextable)
library(officer)


#source_url("https://raw.githubusercontent.com/scottfjennings/HEP_data_work/master/HEP_code/HEP_utility_functions.R")
source("C:/Users/scott.jennings/OneDrive - Audubon Canyon Ranch/Projects/core_monitoring_research/HEP/hep_analyses/how_are_the_egrets_doing/code/ms_analysis/hep_trend_utilities.R")
source("C:/Users/scott.jennings/OneDrive - Audubon Canyon Ranch/Projects/core_monitoring_research/HEP/HEP_data_work/HEP_code/HEP_utility_functions.R")

```

```{r}
# prepare data ----
hepdata_location = here("C:/Users/scott.jennings/OneDrive - Audubon Canyon Ranch/Projects/core_monitoring_research/HEP/HEP_data_work/HEP_data/HEPDATA.accdb")
# all these functions are in HEP_utility_functions.R
hep_sites <- hep_sites_from_access(hepdata_location) %>% 
  dplyr::select(code, parent.code, site.name, parent.site.name, subregion)

options(scipen = 999)

start.year = 1995
end.year = 2019

sfbbo_nests <- read.csv(here("data/CWB_peak_hep_nest_locs_through_2019.csv")) %>% 
  mutate(ColonyName = ifelse(ColonyName == "Lake Merced North" & Latitude == 37.728057, "Lake Merced Mesa", ColonyName)) %>% 
  dplyr::select("site.name" = ColonyName, "species" = SpeciesCode, "year" = SurveyYear, "peakactvnsts" = PeakNumberofNests) %>% 
  full_join(read.csv(here("data/all_sfbbo_sites_subreg.csv"))) %>% 
  filter(!site.name %in% c("Bacon Island", "Elmwood Correctional", "Sandy Wool Lake"))

analysis_table <- readRDS("C:/Users/scott.jennings/OneDrive - Audubon Canyon Ranch/Projects/core_monitoring_research/HEP/HEP_data_work/HEP_data/hep_annual_nest_abundance") %>%
  full_join(hep_sites) %>%
  filter(site.name != "Brentwood") %>% 
  bind_rows(sfbbo_nests) %>% 
  filter(!species %in% c("DCCO", "CAEG"), peakactvnsts >= 0, !is.na(year), between(year, start.year, end.year)) %>% 
  #cut_never_nested() %>% 
  dplyr::select(year, subregion, code, species, peakactvnsts) 

# calculate a cumulative rain index following Stenzel and Page 2018
rain_lag <- readRDS(here("data/subreg_rain")) %>% 
  dplyr::select(year = birdyear, subregion, subreg.name, subreg.rain) %>%
  data.frame() %>% 
  arrange(subregion, year) %>% 
  group_by(subregion) %>% 
  mutate(subreg.rain = subreg.rain + (lag(subreg.rain)/2) + (lag(subreg.rain, 2)/3))
  
subreg_mean_rain_lag <- rain_lag %>% 
  filter(between(year, start.year, end.year)) %>% 
  group_by(subregion) %>% 
  summarise(mean.subreg.rain = mean(subreg.rain))

trend_analysis_table1 <- analysis_table  %>%
  group_by(year, subregion, species) %>% 
  summarise(tot.nests = sum(peakactvnsts)) %>% 
  ungroup() %>% 
  bind_rows(analysis_table %>% # add duplicate of all data with subregion set to "All" 
              group_by(year, species) %>%
              summarise(tot.nests = sum(peakactvnsts)) %>%
              ungroup() %>%
              mutate(subregion = "All")) %>% 
    full_join(expand.grid(species = distinct(analysis_table, species)$species,
                          year = seq(start.year, end.year, by = 1),
                          subregion = subreg_key$subregion)) %>% 
    mutate(tot.nests = ifelse(is.na(tot.nests), 0, tot.nests))


start_end_years <- trend_analysis_table1 %>% 
  filter(tot.nests > 0) %>% 
  group_by(subregion, species) %>% 
  summarise(spp.sub.start.year = min(year) - 1,
            spp.sub.end.year = max(year) + 1) %>% 
  ungroup() %>% 
  mutate(spp.sub.start.year = ifelse(spp.sub.start.year < start.year, start.year, spp.sub.start.year),
         spp.sub.end.year = ifelse(spp.sub.end.year > end.year, end.year, spp.sub.end.year))

spp_subreg_nyears <- trend_analysis_table1 %>%
  filter(tot.nests > 0) %>% 
  group_by(subregion, species) %>% 
  summarise(nyears = n())
  


# for mapping through species and subregions
spp_subreg <- trend_analysis_table1 %>% 
  right_join(spp_subreg_nyears %>% filter(nyears >= 5) %>% dplyr::select(-nyears)) %>% 
  distinct(species, subregion) %>% 
  mutate(spp.subreg = paste(species, subregion, sep = "_"))


trend_analysis_table <- trend_analysis_table1 %>% 
  full_join(start_end_years) %>% 
  mutate(zkeep = ifelse(year >= spp.sub.start.year & year <= spp.sub.end.year, TRUE, FALSE)) %>% 
  arrange(subregion, species, year) %>% 
  filter(zkeep == TRUE) %>% 
  dplyr::select(-zkeep) %>% 
  left_join(rain_lag) 



n_colonies <- analysis_table %>% 
  filter(peakactvnsts > 0, !is.na(species)) %>% 
  group_by(subregion, species, year) %>% 
  summarise(n.colonies = n()) %>% 
  left_join(subreg_key) 


```


```{r}

spp_subreg_mods <- map2(spp_subreg$species, spp_subreg$subregion, fit_mods_glmbn_year2)
names(spp_subreg_mods) <- spp_subreg$spp.subreg

#spp_subreg_mods$SNEG_SUS = fit_mods_glmbn("SNEG", "SUS")


```


```{r}

preds <- map2_df(spp_subreg_mods, names(spp_subreg_mods), get_preds_glmnb)
coef_ci <- map2_df(spp_subreg_mods, names(spp_subreg_mods), get_coefs_cis)

saveRDS(preds, here("output/model_predictions"))
```

```{r}
# from https://stackoverflow.com/questions/54438495/shift-legend-into-empty-facets-of-a-faceted-plot-in-ggplot2
# move legend to free facet spot

shift_legend <- function(p){

  # check if p is a valid object
  if(!"gtable" %in% class(p)){
    if("ggplot" %in% class(p)){
      gp <- ggplotGrob(p) # convert to grob
    } else {
      message("This is neither a ggplot object nor a grob generated from ggplotGrob. Returning original plot.")
      return(p)
    }
  } else {
    gp <- p
  }

  # check for unfilled facet panels
  facet.panels <- grep("^panel", gp[["layout"]][["name"]])
  empty.facet.panels <- sapply(facet.panels, function(i) "zeroGrob" %in% class(gp[["grobs"]][[i]]))
  empty.facet.panels <- facet.panels[empty.facet.panels]
  if(length(empty.facet.panels) == 0){
    message("There are no unfilled facet panels to shift legend into. Returning original plot.")
    return(p)
  }

  # establish extent of unfilled facet panels (including any axis cells in between)
  empty.facet.panels <- gp[["layout"]][empty.facet.panels, ]
  empty.facet.panels <- list(min(empty.facet.panels[["t"]]), min(empty.facet.panels[["l"]]),
                             max(empty.facet.panels[["b"]]), max(empty.facet.panels[["r"]]))
  names(empty.facet.panels) <- c("t", "l", "b", "r")

  # extract legend & copy over to location of unfilled facet panels
  guide.grob <- which(gp[["layout"]][["name"]] == "guide-box")
  if(length(guide.grob) == 0){
    message("There is no legend present. Returning original plot.")
    return(p)
  }
  gp <- gtable_add_grob(x = gp,
                        grobs = gp[["grobs"]][[guide.grob]],
                        t = empty.facet.panels[["t"]],
                        l = empty.facet.panels[["l"]],
                        b = empty.facet.panels[["b"]],
                        r = empty.facet.panels[["r"]],
                        name = "new-guide-box")

  # squash the original guide box's row / column (whichever applicable)
  # & empty its cell
  guide.grob <- gp[["layout"]][guide.grob, ]
  if(guide.grob[["l"]] == guide.grob[["r"]]){
    gp <- gtable_squash_cols(gp, cols = guide.grob[["l"]])
  }
  if(guide.grob[["t"]] == guide.grob[["b"]]){
    gp <- gtable_squash_rows(gp, rows = guide.grob[["t"]])
  }
  gp <- gtable_remove_grobs(gp, "guide-box")

  return(gp)
}

```


```{r}


plot_pred_comparison <- function(zspp) {
  raw_dat <- trend_analysis_table1 %>% 
    filter(species == zspp) %>% 
    dplyr::select(species, subregion, year, tot.nests)


  
  pred_dat <- preds %>% 
    filter(species == zspp) %>% 
    full_join(raw_dat) %>% 
    full_join(subreg_key) %>% 
    group_by(subregion, species) %>% 
    mutate(text.y = ceiling(max(tot.nests) * 1.1))

  
mod_pred_plot <- pred_dat %>% 
  ggplot() +
  geom_line(aes(x = year, y = estimate), size = 0.5) +
  geom_ribbon(aes(x = year, ymin = lci, ymax = uci), alpha = 0.25) +
  geom_point(aes(x = year, y = tot.nests), size = 0.5)  + 
  scale_x_continuous(breaks = seq(1995, 2020, by = 5), labels = seq(1995, 2020, by = 5)) + 
  scale_y_continuous(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.3))))) +
  facet_wrap(~subreg.name, scales = "free_y", labeller = labeller(subreg.name = label_wrap_gen(30))) +
  labs(y = "Nest abundance",
       x = "Year",
       title = "") +
  theme_bw() +
  theme(text = element_text(size=8))

ggsave(paste("C:/Users/scott.jennings/Documents/Projects/core_monitoring_research/HEP/hep_analyses/how_are_the_egrets_doing/figures/mod_pred_plot_", zspp, ".png", sep = ""), mod_pred_plot, width = 7, height = 5, dpi = 600)

mod_pred_plot

}

```


```{r}
pred_table_start <- preds %>% 
  group_by(species, subregion) %>% 
  mutate(keep.year = case_when(year == min(year) ~ "first.year",
                               year == max(year) ~ "last.year",
                               estimate == max(estimate) ~ "maxest.year")) %>% 
  ungroup() %>% 
  filter(!is.na(keep.year)) %>% 
  mutate(across(c(estimate, lci, uci), ~round(., 0)),
         out.estimate = paste(estimate, " (", lci, "-", uci, ")", sep = "")) 

per_change <- pred_table_start  %>% 
  filter(keep.year != "maxest.year") %>% 
  pivot_wider(id_cols = c(species, subregion), names_from = keep.year, values_from = estimate) %>% 
  mutate(abs.change = last.year - first.year,
         per.change = ((last.year/first.year) - 1) * 100,
         per.change = round(per.change, 1)) %>% 
  dplyr::select(species, subregion, per.change, abs.change)
  

pred_table_est <- pred_table_start %>% 
  mutate(keep.year = paste("est.", keep.year, sep = "")) %>% 
  pivot_wider(id_cols = c(species, subregion), names_from = keep.year, values_from = out.estimate)

pred_table_years <- pred_table_start  %>% 
  mutate(year = as.character(year))%>% 
  pivot_wider(id_cols = c(species, subregion), names_from = keep.year, values_from = year)

pred_table <- full_join(pred_table_est, pred_table_years) %>% 
  full_join(subreg_key) %>% 
  full_join(per_change) %>% 
  arrange(species, subreg.name)

write.csv(pred_table, here("output/trend_model_predictions.csv"), row.names = FALSE)

```

```{r}

max_min_per_change <- pred_table %>%
  filter(per.change != Inf & per.change != -Inf, subregion != "All") %>%
  dplyr::select(species, subreg.name, contains("first.year"), contains("last.year"), per.change) %>%  
  group_by(species) %>% 
  mutate(max.per.change = per.change == max(per.change),
         min.per.change = per.change == min(per.change)) %>% 
  ungroup() %>% 
  filter(max.per.change == TRUE | min.per.change == TRUE)

max_min_abs_change <- pred_table %>%
filter(subregion != "All") %>% 
  dplyr::select(species, subreg.name, contains("first.year"), contains("last.year"), abs.change) %>%  
  group_by(species) %>% 
  mutate(max.abs.change = abs.change == max(abs.change),
         min.abs.change = abs.change == min(abs.change)) %>% 
  ungroup() %>% 
  filter(max.abs.change == TRUE | min.abs.change == TRUE)



make_change_text <- function(zspp) {
  per_increase_spp <- filter(max_min_per_change, species == zspp, max.per.change == TRUE)
  abs_increase_spp <- filter(max_min_abs_change, species == zspp, max.abs.change == TRUE)
  per_decrease_spp <- filter(max_min_per_change, species == zspp, min.per.change == TRUE)
  abs_decrease_spp <- filter(max_min_abs_change, species == zspp, min.abs.change == TRUE)
  

  change_text <- paste("For ", translate_bird_names(zspp, "alpha.code", "common.name"), 
  " (Table XXX), the largest percent increase was observed in ", per_increase_spp$subreg.name, " where abundance went from ",
  per_increase_spp$est.first.year, " nests in ", per_increase_spp$first.year, 
  " to ", per_increase_spp$est.last.year, " nests in ", per_increase_spp$last.year, ". The largest absolute increase was observed in ",
  abs_increase_spp$subreg.name, " where abundance increased by ", abs_increase_spp$abs.change, " nests.", 
  # decrease
  " The largest percent decrease in ", translate_bird_names(zspp, "alpha.code", "common.name"), " nest abundance was in ",
  per_decrease_spp$subreg.name, " where nesting abundance decreased from ",
  per_decrease_spp$est.first.year,  " nests in ", per_decrease_spp$first.year, 
  " to ", per_decrease_spp$est.last.year, " nests in ", per_decrease_spp$last.year, ". The largest absolute decrease was observed in ",
  abs_decrease_spp$subreg.name, " where abundance dropped by ", -1 * abs_decrease_spp$abs.change, " nests.",
  sep = "")
  
 return(change_text) 
  
}
```





```{r}
make_spp_tab <- function(zspp) {
  ztable <- filter(pred_table, species == zspp) %>% 
    full_join(subreg_key %>% dplyr::select(subreg.name)) %>% 
    arrange(subreg.name) %>% 
    mutate(break1 = "",
           subreg.name = ifelse(is.na(first.year), paste(subreg.name, "*", sep = ""), as.character(subreg.name))) %>% 
    dplyr::select(subreg.name, first.year, est.first.year, break1, last.year, est.last.year, per.change, -species)
  
  #zfoots <- sprintf("%.0f", which(is.na(ztable$first.year)))
  
    flextable(ztable) %>%
    add_header_row(values = c("", "First year observed", "", 
                              "Final year observed", ""), colwidths = c(1, 2, 1, 2, 1)) %>%
    set_header_labels(subreg.name = "Subregion",
                    first.year = "Year",
                    est.first.year = "Estimate (95% CI)",
                    break1 = "",
                    last.year = "Year",
                    est.last.year = "Estimate (95% CI)",
                    per.change = "% change") %>%
    align(j = 2:7, align = "center", part = "all") %>%
    border_remove() %>%
    border(i = 1, border.top = fp_border(color = "black"), part = "header") %>%
    #border(i = 2, border.bottom = fp_border(color = "black"), part = "header") %>%
    #hline_top(part = "body") %>%
    hline_bottom(part = "body") %>%
    fit_to_width(max_width = 8.5) %>% 
    width(j = c(1, 3, 6), width = 1.5) %>% 
    width(j = c(2, 5), width = .5) %>% 
    width(j = 7, width = 1) %>% 
      width(j = 4, width = 0.1) %>% 
     # footnote(i = 1, j = 1, value = as_paragraph(c("Data insufficient to model")), part = "body") %>% 
    set_caption(caption = paste("Changes in number of observed ", translate_bird_names(zspp, "alpha.code", "common.name"), " nests by subregion in the San Francisco Bay area, 1995-2019. Shown for each subregion are the first and last years a species was observed nesting, the estimated nest abundance and 95% Confidence interval, and the percent change from the first to final year."))
}

```

`r make_change_text("GREG")`

```{r}
make_spp_tab("GREG")
```

<br>  
<br>  

`r make_change_text("GBHE")`

<br>
```{r}
make_spp_tab("GBHE")
```

<br>  
<br>  

`r make_change_text("SNEG")`

<br>

```{r}
make_spp_tab("SNEG")
```

<br>  
<br>  


`r make_change_text("BCNH")` 

<br>

```{r}
make_spp_tab("BCNH")
```


Figures

Figure XXX. Number of active (≥1 nest) heron and egret colonies monitored in each subregion in the San Francisco Bay Area, 1995-2019. 
```{r}


  num_cols_plot <- ggplot(n_colonies) +
  geom_line(aes(x = year, y = n.colonies, color = translate_bird_names(species, "alpha.code", "common.name")))  + 
  scale_x_continuous(breaks = seq(1995, 2020, by = 5), labels = seq(1995, 2020, by = 5)) + 
  scale_y_continuous(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.1))))) +
  facet_wrap(~subreg.name, scales = "free_y", labeller = labeller(subreg.name = label_wrap_gen(30))) +
  labs(y = "Number of colonies with \u2265 1 nest",
       x = "Year",
       color = "") +
  theme_bw() +
  theme(text = element_text(size=8))

grid.draw(shift_legend(num_cols_plot))


ggsave("C:/Users/scott.jennings/Documents/Projects/core_monitoring_research/HEP/hep_analyses/how_are_the_egrets_doing/figures/n_colonies.png", width = 7, height = 5, dpi = 300)

```

<br>  
<br>  

Figure XXX. Changes in nest abundance of Great Egret by subregion in the San Francisco Bay Area, 1995-2019. Shown are the estimated mean abundance (line) and 95% Confidence interval (gray ribbon) from a generalize linear model and the raw data (dots) the models were fitted to.

```{r}

plot_pred_comparison("GREG")

```

<br>  
<br>  


Figure XXX. Changes in nest abundance of Great Blue Heron by subregion in the San Francisco Bay Area, 1995-2019. Shown are the estimated mean abundance (line) and 95% Confidence interval (gray ribbon) from a generalize linear model and the raw data (dots) the models were fitted to.

```{r}

plot_pred_comparison("GBHE")

```


<br>  
<br>  

Figure XXX. Changes in nest abundance of Snowy Egret by subregion in the San Francisco Bay Area, 1995-2019. Shown are the estimated mean abundance (line) and 95% Confidence interval (gray ribbon) from a generalize linear model and the raw data (dots) the models were fitted to.
```{r}

plot_pred_comparison("SNEG")

```

<br>  
<br>  


Figure XXX. Changes in nest abundance of Black-crowned Night-Heron by subregion in the San Francisco Bay Area, 1995-2019. Shown are the estimated mean abundance (line) and 95% Confidence interval (gray ribbon) from a generalize linear model and the raw data (dots) the models were fitted to.
```{r}

plot_pred_comparison("BCNH")

```



Figure XXX. Absolute and percent changes in the number of observed heron and egret nests by subregion in the San Francisco Bay area, 1995-2019. Shown for each subregion are the change from the first to last years a species was observed nesting. All estimates are extracted from the fit of a genralized linear model.
```{r}
change_mag <- pred_table %>% 
  dplyr:: select(subreg.name, species, contains("change")) %>% 
  full_join(expand.grid(species = distinct(pred_table, species)$species,
                        subreg.name = distinct(pred_table, subreg.name)$subreg.name)) %>% 
  pivot_longer(cols = contains("change"), names_to = "change.type", values_to = "change") %>% 
  mutate(change = ifelse(is.na(change), 0, change),
         species = translate_bird_names(species, "alpha.code", "common.name"),
         change.type = ifelse(change.type == "abs.change", "Absolute change", "Percent change")) %>% 
  ggplot(aes(x = species, y = change, fill = species))  + 
  geom_bar(position="dodge", stat="identity") +
  geom_hline(yintercept = 0) +
  facet_grid(subreg.name~change.type, scales = "free", labeller = labeller(subreg.name = label_wrap_gen(30))) +
  labs(x = "",
       y = "Change magnitude",
       fill = "") +
  theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        strip.text.y.right = element_text(angle = 0))


ggsave("C:/Users/scott.jennings/Documents/Projects/core_monitoring_research/HEP/hep_analyses/how_are_the_egrets_doing/figures/change_magnitude.png", width = 7, height = 7, dpi = 300)

change_mag
```



```{r}
pred_table_start %>% 
  filter(keep.year %in% c("first.year", "last.year")) %>% 
  dplyr::select(subregion, species, keep.year, estimate) %>% 
  pivot_wider(names_from = keep.year, values_from = estimate) %>% 
  mutate(change = last.year - first.year,
         out.change = paste(change, " (", first.year, ", ", last.year, ")", sep = ""),
         species = translate_bird_names(species, "alpha.code", "common.name")) %>% 
  full_join(subreg_key) %>%
  dplyr::select(subreg.name, species, out.change) %>% 
  pivot_wider(names_from = species, values_from = out.change) %>% 
  arrange(subreg.name) %>%
  dplyr::select("subreg.name", "Great Egret", "Great Blue Heron", "Snowy Egret", "Black-crowned Night-Heron") %>% 
  flextable() %>% 
  set_header_labels(subreg.name = "") %>% 
  width(j = 1, width = 1.75) %>%  
  width(j = 2:5, width = 1.25) %>% 
  fit_to_width(max_width = 7.5) %>% 
    set_caption(caption = paste("Changes in absolute number of observed heron and egret nests by subregion in the San Francisco Bay area, 1995-2019. Shown for each subregion are the change in number of nests from the first to last years a species was observed nesting, and in parentheses the values in the first and last years. All estimates are extracted from the fit of a genralized linear model."))

```




```{r}
pred_table_start %>% 
  filter(keep.year %in% c("first.year", "last.year")) %>% 
  dplyr::select(subregion, species, keep.year, estimate) %>% 
  pivot_wider(names_from = keep.year, values_from = estimate) %>% 
  mutate(change = ((last.year/first.year) - 1) * 100,
         out.change = paste(round(change, 1), " (", first.year, ", ", last.year, ")", sep = ""),
         species = translate_bird_names(species, "alpha.code", "common.name")) %>% 
  full_join(subreg_key) %>%
  dplyr::select(subreg.name, species, out.change) %>% 
  pivot_wider(names_from = species, values_from = out.change) %>% 
  arrange(subreg.name) %>%
  dplyr::select("subreg.name", "Great Egret", "Great Blue Heron", "Snowy Egret", "Black-crowned Night-Heron") %>% 
  flextable() %>% 
  set_header_labels(subreg.name = "") %>% 
  width(j = 1, width = 1.75) %>%  
  width(j = 2:5, width = 1.25) %>% 
  fit_to_width(max_width = 7.5) %>% 
    set_caption(caption = paste("Changes in number of observed heron and egret nests by subregion in the San Francisco Bay area, 1995-2019. Shown for each subregion are the percent change in number of nests from the first to last years a species was observed nesting, and in parentheses the values in the first and last years. All estimates are extracted from the fit of a genralized linear model."))

```



Figure xxx. Effect of rainfall (previous 3 years cumulative) on number of observed heron and egret nests in each subregion in the San Francisco Bay Area, 1995-2019. 
```{r}
rain_effect <- coef_ci %>% 
filter(varb == "subreg.rain", !is.na(coef)) %>% 
  full_join(expand.grid(species = distinct(coef_ci, species)$species,
                        subregion = distinct(coef_ci, subregion)$subregion)) %>% 
  full_join(subreg_key) %>% 
  mutate(species = translate_bird_names(species, "alpha.code", "common.name"),
         coef = ifelse(is.na(coef), 0, coef)) %>%
ggplot(aes(x = species, y = coef, fill = species))  + 
  geom_bar(position="dodge", stat="identity") +
  facet_wrap(~subreg.name, scales = "free", labeller = labeller(subreg.name = label_wrap_gen(30))) +
  labs(x = "",
       y = "Rain effect",
       fill = "") +
  theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        strip.text.y.right = element_text(angle = 0))+
  theme(text = element_text(size=8))



ggsave("C:/Users/scott.jennings/Documents/Projects/core_monitoring_research/HEP/hep_analyses/how_are_the_egrets_doing/figures/rain_effect.png", width = 7, height = 7, dpi = 300)



grid.draw(shift_legend(rain_effect))

```



Results of sensitivity analysis

```{r}

make_sensitivity_plot <- function(zspp) {
   pred_dat <- preds %>%  
    full_join(subreg_key) %>% 
    filter(species == zspp, subregion != "All")
   
   bootz <- readRDS("C:/Users/scott.jennings/Documents/Projects/core_monitoring_research/HEP/hep_analyses/how_are_the_egrets_doing/output/sensitivity_bootz") %>% 
     full_join(subreg_key) %>% 
     filter(species == zspp, subregion != "All")
   
   
  
sensitivity_plot <- pred_dat %>% 
  ggplot() +
  geom_line(data = pred_dat, aes(x = year, y = estimate), size = 0.5) +
  geom_ribbon(data = bootz, aes(x = year, ymin = lci, ymax = uci), alpha = 0.25) +
  scale_x_continuous(breaks = seq(1995, 2020, by = 5), labels = seq(1995, 2020, by = 5)) + 
  #scale_y_continuous(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.3))))) +
  facet_wrap(~subreg.name, scales = "free_y", labeller = labeller(subreg.name = label_wrap_gen(30))) +
  labs(y = "Nest abundance",
       x = "Year",
       title = "") +
  theme_bw() +
  theme(text = element_text(size=8))

ggsave(paste("C:/Users/scott.jennings/Documents/Projects/core_monitoring_research/HEP/hep_analyses/how_are_the_egrets_doing/figures/sensitivity_plot_", zspp, ".png", sep = ""), sensitivity_plot, width = 7, height = 5, dpi = 300)

sensitivity_plot

}


```

Figure XXX. Results of analysis testing sensitivity of estimates of Great Egret nesting abundance to missing data, by subregion in the San Francisco Bay Area, 1995-2019. The analysis consisted of randomly removing 5% of the annual colony-level observations, refitting the model used in the main analysis, and extracting model predictions as in the main analysis. This process was repeated 1000 times, and 95% confidence intervals for the bootstrapped model predictions were generated. These 95% CI are shown in gray, and the main analysis model predictions (using all the data) are shown by the black line.
```{r}
make_sensitivity_plot("GREG")

```


Figure XXX. Results of analysis testing sensitivity of estimates of Great Blue Heron nesting abundance to missing data, by subregion in the San Francisco Bay Area, 1995-2019. The analysis consisted of randomly removing 5% of the annual colony-level observations, refitting the model used in the main analysis, and extracting model predictions as in the main analysis. This process was repeated 1000 times, and 95% confidence intervals for the bootstrapped model predictions were generated. These 95% CI are shown in gray, and the main analysis model predictions (using all the data) are shown by the black line.
```{r}


make_sensitivity_plot("GBHE")

```



Figure XXX. 

```{r}


plot_pred_comparison <- function(zspp) {
  raw_dat <- trend_analysis_table1 %>% 
    filter(subregion == "All") %>% 
    dplyr::select(species, subregion, year, tot.nests)


  
  pred_dat <- preds %>% 
    filter(subregion == "All") %>% 
    full_join(raw_dat) %>% 
    full_join(subreg_key) %>% 
    group_by(subregion, species) %>% 
    mutate(text.y = ceiling(max(tot.nests) * 1.1)) %>% 
    ungroup() %>% 
    filter(!is.na(species)) %>% 
    mutate(species = translate_bird_names(species, "alpha.code", "common.name"))

  
mod_pred_plot <- pred_dat %>% 
  ggplot() +
  geom_line(aes(x = year, y = estimate), size = 0.5) +
  geom_ribbon(aes(x = year, ymin = lci, ymax = uci), alpha = 0.25) +
  geom_point(aes(x = year, y = tot.nests), size = 0.5)  + 
  #scale_x_continuous(breaks = seq(1995, 2020, by = 5), labels = seq(1995, 2020, by = 5)) + 
  #scale_y_continuous(breaks = function(x) unique(floor(pretty(seq(0, (max(x) + 1) * 1.3))))) +
  facet_wrap(~species, scales = "free_y") +
  labs(y = "Nest abundance",
       x = "Year",
       title = "") +
  theme_bw()

ggsave("C:/Users/scott.jennings/Documents/Projects/core_monitoring_research/HEP/hep_analyses/how_are_the_egrets_doing/figures/mod_pred_plot_allspp_entireregion.png", mod_pred_plot, width = 7, height = 5, dpi = 600)

mod_pred_plot

}

```



