---
title: "Changes table"
output: word_document
date: "`r Sys.Date()`"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      fig.width = 7,
                      fig.height = 9)
```



```{r}
library(tidyverse)
library(birdnames)
library(here)
library(flextable)
library(ftExtra)
library(officer)

#source_url("https://raw.githubusercontent.com/scottfjennings/HEP_data_work/master/HEP_code/HEP_utility_functions.R")
source("C:/Users/scott.jennings/OneDrive - Audubon Canyon Ranch/Projects/core_monitoring_research/HEP/hep_analyses/how_are_the_egrets_doing/code/ms_analysis/hep_trend_utilities.R")
source("C:/Users/scott.jennings/OneDrive - Audubon Canyon Ranch/Projects/core_monitoring_research/HEP/HEP_data_work/HEP_code/HEP_utility_functions.R")

```

```{r}

all_mod_av_preds <- readRDS(here("fitted_models/all_mod_av_preds"))

trend_analysis_table <- readRDS(here("data/trend_analysis_table"))


all_mod_av_preds_per_change <- readRDS(here("fitted_models/all_mod_av_per_change"))

```



```{r}
pred_table_start <- all_mod_av_preds %>% 
  group_by(species, subregion) %>% 
  mutate(keep.year = case_when(year == min(year) ~ "first.year",
                               year == max(year) ~ "last.year")) %>% 
  ungroup() %>% 
  filter(!is.na(keep.year)) %>% 
  mutate(across(c(mod.avg.pred, lower.CL, upper.CL), ~round(., 0)),
         out.estimate = paste(mod.avg.pred, " (", lower.CL, ", ", upper.CL, ")", sep = "")) 

#per_change <- pred_table_start  %>% 
#  filter(keep.year != "maxest.year") %>% 
#  pivot_wider(id_cols = c(species, subregion), names_from = keep.year, values_from = mod.avg.pred) %>% 
#  mutate(abs.change = last.year - first.year,
#         per.change = ((last.year/first.year) - 1) * 100,
#         per.change = round(per.change, 1)) %>% 
#  dplyr::select(species, subregion, per.change, abs.change)
 
per_change <- all_mod_av_preds_per_change %>% 
  separate(alpha.code, c("species", "subregion"), sep = "_") %>% 
  mutate(across(c(per.change, per.change.LCI, per.change.UCI, marginal.effect, marginal.lower.CL, marginal.upper.CL), ~round(., 0)),
         out.per.change = paste(per.change, "% (", per.change.LCI, "%, ", per.change.UCI, "%)", sep = ""),
         out.abs.change = paste(marginal.effect, " (", marginal.lower.CL, ", ", marginal.upper.CL, ")", sep = "")) %>%  
  dplyr::select(species, subregion, out.per.change, out.abs.change)
  

pred_table_est <- pred_table_start %>% 
  mutate(keep.year = paste("est.", keep.year, sep = "")) %>% 
  pivot_wider(id_cols = c(species, subregion), names_from = keep.year, values_from = out.estimate)

pred_table_years <- pred_table_start  %>% 
  mutate(year = as.character(year))%>% 
  pivot_wider(id_cols = c(species, subregion), names_from = keep.year, values_from = year)

pred_table <- full_join(pred_table_est, pred_table_years) %>% 
  full_join(subreg_key) %>% 
  full_join(per_change) %>% 
  arrange(species, subreg.name)

write.csv(pred_table, here("output/trend_model_predictions.csv"), row.names = FALSE)



```




```{r}


spp_subreg <- readRDS(here("data/spp_subreg")) %>%
  mutate(species = factor(species, levels = c("GREG", "GBHE", "SNEG", "BCNH"))) %>% 
  full_join(subreg_key) %>% 
  arrange(species, subreg.name) 


pred_table_out <-  pred_table %>% 
    #left_join(readRDS(here("fitted_models/change_evidence"))) %>% 
    arrange(subreg.name) %>% 
  mutate(common.name = translate_bird_names(species, "alpha.code", "common.name"),
         species = factor(species, levels = levels(spp_subreg$species))) %>%  
  arrange(species, subreg.name) %>% 
    mutate(break1 = "",
           subreg.name = ifelse(is.na(first.year), paste(subreg.name, "*", sep = ""), as.character(subreg.name)))
  
# pred_table_out %>% group_by(species) %>% filter(subregion != "All", spp.subreg != "SNEG_SUS") %>% filter(abs.change == max(abs.change) | abs.change == min(abs.change)) %>% view()

  #zfoots <- sprintf("%.0f", which(is.na(ztable$first.year)))
```

```{r eval=FALSE}

# Table XX. Changes in estimated number of nests of four Ardeid species by subregion in the San Francisco Bay area, 1995-2019. Shown for each species and subregion are the first and last years a species was observed nesting, the estimated nest abundance and 95% Confidence interval in the first and last years, and the percent change from the first to last year. An * following a subregion name indicates there was insufficient data for modeling. An * following percent change signifies no models with year structure were competitive (∆AICc ≤ 2) and/or the intercept only model was competitive, indicating poor statistical evidence for any change over time.
    pred_table_out %>%
      mutate(common.name = ifelse(subregion == "All", common.name, "")) %>% 
    dplyr::select(common.name, subreg.name, first.year, est.first.year, break1, last.year, est.last.year, out.per.change, -species) %>%
      flextable() %>% 
      add_header_row(values = c("", "First year observed", "", 
                              "Final year observed", ""), colwidths = c(2, 2, 1, 2, 1)) %>%
    set_header_labels(common.name = "Species",
                      subreg.name = "Subregion",
                    first.year = "Year",
                    est.first.year = "Estimate (95% CI)",
                    break1 = "",
                    last.year = "Year",
                    est.last.year = "Estimate (95% CI)",
                    out.per.change = "% change") %>%
    align(j = 3:8, align = "center", part = "all") %>% 
  fit_to_width(max_width = 7.5) %>%
  border(i = c(9, 20, 28), part = "body", border.bottom = fp_border(color = "black")) %>% 
  colformat_md() %>%
    fit_to_width(max_width = 8.5) %>% 
    width(j = 1, width = 1) %>% 
    width(j = c(2, 4, 7), width = 1.5) %>% 
    width(j = c(3, 6), width = .5) %>% 
    width(j = 8, width = 1) %>% 
      width(j = 5, width = 0.1) 


```





```{r}
make_spp_tab <- function(zspp) {
  ztable <- filter(pred_table, species == zspp) %>% 
    full_join(subreg_key %>% dplyr::select(subreg.name))  %>% 
    #left_join(readRDS(here("fitted_models/change_evidence"))) %>% 
    arrange(subreg.name) %>% 
    mutate(break1 = "",
           subreg.name = ifelse(is.na(first.year), paste(subreg.name, "*", sep = ""), as.character(subreg.name))) %>% 
    dplyr::select(subreg.name, first.year, est.first.year, break1, last.year, est.last.year, out.per.change, -species)
  
  #zfoots <- sprintf("%.0f", which(is.na(ztable$first.year)))
  
    flextable(ztable) %>%
    add_header_row(values = c("", "First year observed", "", 
                              "Final year observed", ""), colwidths = c(1, 2, 1, 2, 1)) %>%
    set_header_labels(subreg.name = "Subregion",
                    first.year = "Year",
                    est.first.year = "Estimate (95% CI)",
                    break1 = "",
                    last.year = "Year",
                    est.last.year = "Estimate (95% CI)",
                    out.per.change = "% change (95% CI)") %>%
    align(j = 2:7, align = "center", part = "all") %>%
    border_remove() %>%
    border(i = 1, border.top = fp_border(color = "black"), part = "header") %>%
    #border(i = 2, border.bottom = fp_border(color = "black"), part = "header") %>%
    #hline_top(part = "body") %>%
    hline_bottom(part = "body") %>%
    fit_to_width(max_width = 8.5) %>% 
    width(j = c(1, 3, 6, 7), width = 1.5) %>% 
    width(j = c(2, 5), width = .5) %>% 
    width(j = 7, width = 1) %>% 
      width(j = 4, width = 0.1) 
}

make_spp_tab2 <- function(zspp) {
  ztable <- filter(pred_table, species == zspp) %>% 
    full_join(subreg_key %>% dplyr::select(subreg.name))  %>% 
    arrange(subreg.name) %>% 
    mutate(years.obs = paste(first.year, last.year, sep = ", "),
           break1 = "",
           subreg.name = ifelse(is.na(first.year), paste(subreg.name, "*", sep = ""), as.character(subreg.name))) %>% 
    dplyr::select(subreg.name, years.obs, out.abs.change, out.per.change, est.last.year)
  
  
    flextable(ztable) %>%
    set_header_labels(subreg.name = "Subregion",
                    years.obs = "First and last\nyear observed",
                    est.first.year = "Estimate (95% CI)",
                    est.last.year = "Final year estimated\nabundance (95% CI)",
                    out.per.change = "% change (95% CI)",
                    out.abs.change = "Absolute change (95% CI)") %>%
    align(j = 2:5, align = "center", part = "all") %>%
    border_remove() %>%
    border(i = 1, border.top = fp_border(color = "black"), part = "header") %>%
    hline_bottom(part = "body") %>%
    fit_to_width(max_width = 8.5) %>% 
    width(j = 1:5, width = 1.5)
}

```



```{r eval=FALSE}

# get the min and max percent change for each species
max_min_change <- all_mod_av_preds_per_change %>% 
  separate(alpha.code, c("species", "subregion"), sep = "_") %>%
  filter(per.change != Inf & per.change != -Inf, subregion != "All") %>%
  left_join(subreg_key) %>% 
  dplyr::select(species, subreg.name, contains("first.year"), contains("last.year"), per.change) %>%  
  group_by(species) %>% 
  mutate(max.per.change = per.change == max(per.change),
         min.per.change = per.change == min(per.change)) %>% 
  ungroup() %>% 
  filter(max.per.change == TRUE | min.per.change == TRUE)

max_min_abs_change <- all_mod_av_preds_per_change %>% 
  separate(alpha.code, c("species", "subregion"), sep = "_") %>%
  filter(per.change != Inf & per.change != -Inf, subregion != "All") %>%
  left_join(subreg_key) %>%   
  dplyr::select(species, subreg.name, contains("first.year"), contains("last.year"), marginal.effect) %>% 
  group_by(species) %>% 
  mutate(max.abs.change = marginal.effect == max(marginal.effect),
         min.abs.change = marginal.effect == min(marginal.effect)) %>% 
  ungroup() %>% 
  filter(max.abs.change == TRUE | min.abs.change == TRUE)



make_change_text <- function(zspp) {
  per_increase_spp <- filter(max_min_per_change, species == zspp, max.per.change == TRUE)
  abs_increase_spp <- filter(max_min_abs_change, species == zspp, max.abs.change == TRUE)
  per_decrease_spp <- filter(max_min_per_change, species == zspp, min.per.change == TRUE)
  abs_decrease_spp <- filter(max_min_abs_change, species == zspp, min.abs.change == TRUE)
  

  change_text <- paste("For ", translate_bird_names(zspp, "alpha.code", "common.name"), 
  " (Table XXX), the largest percent increase was observed in ", per_increase_spp$subreg.name, " where abundance went from ",
  per_increase_spp$est.first.year, " nests in ", per_increase_spp$first.year, 
  " to ", per_increase_spp$est.last.year, " nests in ", per_increase_spp$last.year, ". The largest absolute increase was observed in ",
  abs_increase_spp$subreg.name, " where abundance increased by ", abs_increase_spp$abs.change, " nests.", 
  # decrease
  " The largest percent decrease in ", translate_bird_names(zspp, "alpha.code", "common.name"), " nest abundance was in ",
  per_decrease_spp$subreg.name, " where nesting abundance decreased from ",
  per_decrease_spp$est.first.year,  " nests in ", per_decrease_spp$first.year, 
  " to ", per_decrease_spp$est.last.year, " nests in ", per_decrease_spp$last.year, ". The largest absolute decrease was observed in ",
  abs_decrease_spp$subreg.name, " where abundance dropped by ", -1 * abs_decrease_spp$abs.change, " nests.",
  sep = "")
  
 return(change_text) 
  
}
```




<br>  


<br>  
Table XX. Changes in estimated number of Great Egret nests by subregion in the San Francisco Bay area, 1995-2019. Shown for each subregion are the first and last years a species was observed nesting, the estimated nest abundance and 95% Confidence interval in the first and last years, and the percent change from the first to last year. An * following a subregion name indicates there was insufficient data for modeling. An * following percent change signifies no models with year structure were competitive (∆AICc ≤ 2) and/or the intercept only model was competitive, indicating poor statistical evidence for any change over time.
<br>
```{r}
#`r make_change_text("GREG")`
make_spp_tab("GREG")
```
<br>
```{r}
make_spp_tab2("GREG")
```

<br>  
<br>  

<br>  


<br>  
Table XX. Changes in estimated number of Great Blue Heron nests by subregion in the San Francisco Bay area, 1995-2019. Shown for each subregion are the first and last years a species was observed nesting, the estimated nest abundance and 95% Confidence interval in the first and last years, and the percent change from the first to last year. An * following a subregion name indicates there was insufficient data for modeling. An * following percent change signifies no models with year structure were competitive (∆AICc ≤ 2) and/or the intercept only model was competitive, indicating poor statistical evidence for any change over time.
<br>
```{r}
#`r make_change_text("GBHE")`
make_spp_tab("GBHE")
```
<br>
```{r}
make_spp_tab2("GBHE")
```

<br>  
<br>  

<br>  


<br>  
Table XX. Changes in estimated number of Snowy Egret nests by subregion in the San Francisco Bay area, 1995-2019. Shown for each subregion are the first and last years a species was observed nesting, the estimated nest abundance and 95% Confidence interval in the first and last years, and the percent change from the first to last year. An * following a subregion name indicates there was insufficient data for modeling. An * following percent change signifies no models with year structure were competitive (∆AICc ≤ 2) and/or the intercept only model was competitive, indicating poor statistical evidence for any change over time.
<br>
```{r}
#`r make_change_text("SNEG")`
make_spp_tab("SNEG")
```
<br>
```{r}
make_spp_tab2("SNEG")
```

<br>  
<br>  

 
<br>  


<br>  
Table XX. Changes in estimated number of Black-crowned Night-Heron nests by subregion in the San Francisco Bay area, 1995-2019. Shown for each subregion are the first and last years a species was observed nesting, the estimated nest abundance and 95% Confidence interval in the first and last years, and the percent change from the first to last year. An * following a subregion name indicates there was insufficient data for modeling. An * following percent change signifies no models with year structure were competitive (∆AICc ≤ 2) and/or the intercept only model was competitive, indicating poor statistical evidence for any change over time.
<br>
```{r}
#`r make_change_text("BCNH")`
make_spp_tab("BCNH")
```
<br>
```{r}
make_spp_tab2("BCNH")
```
